---
title: "Sarasota Weather"
subtitle: "The last seven complete days in Sarasota, Florida"
author: ""
format:
  html:
    toc: false
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(lubridate)

weather <- read_csv("data/sarasota-weather.csv", show_col_types = FALSE) |>
  mutate(
    date = as.Date(date),
    day = format(date, "%a %b %-d"),
    weather = factor(weather)
  )

period_start <- min(weather$date)
period_end <- max(weather$date)

summary_stats <- weather |>
  summarise(
    high = max(temperature_f),
    low = min(temperature_f),
    rain = sum(precipitation_in),
    windy = max(wind_mph)
  )
```

## A week in Sarasota

`r format(period_start, "%B %-d")` to `r format(period_end, "%B %-d, %Y")`  
Hourly observations from Open-Meteo, shown in local time.

```{r}
#| label: weather-plot
#| fig-width: 11
#| fig-height: 7
#| fig-cap: "Hourly temperature, humidity, rainfall, and wind speed."

plot_data <- weather |>
  mutate(
    rainfall = if_else(precipitation_in > 0, precipitation_in, NA_real_),
    wind_label = if_else(wind_mph >= 15, sprintf("%.0f mph", wind_mph), NA_character_)
  )

ggplot(plot_data, aes(x = time)) +
  geom_col(aes(y = precipitation_in * 70), fill = "#79c7d3", alpha = 0.55, width = 0.035) +
  geom_ribbon(aes(ymin = temperature_f - 2, ymax = temperature_f + 2),
              fill = "#f3a261", alpha = 0.16) +
  geom_line(aes(y = temperature_f, color = "Temperature"), linewidth = 1.1) +
  geom_line(aes(y = humidity_percent, color = "Humidity"), linewidth = 0.8, alpha = 0.8) +
  geom_point(data = filter(plot_data, wind_mph >= 15),
             aes(y = temperature_f, size = wind_mph), color = "#264653", alpha = 0.75) +
  scale_color_manual(values = c("Temperature" = "#e76f51", "Humidity" = "#277da1")) +
  scale_size_continuous(range = c(2, 5), guide = "none") +
  scale_y_continuous(
    name = "Temperature (F) / relative humidity (%)",
    limits = c(0, 100),
    sec.axis = sec_axis(~ . / 70, name = "Precipitation (inches)")
  ) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%a\n%b %-d", expand = expansion(mult = c(0.01, 0.02))) +
  labs(color = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    plot.background = element_rect(fill = "#fffaf3", color = NA),
    panel.background = element_rect(fill = "#fffaf3", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title = element_text(color = "#264653", face = "bold"),
    axis.text = element_text(color = "#4a5759"),
    legend.position = "top",
    legend.text = element_text(color = "#264653", face = "bold"),
    plot.caption = element_text(color = "#6b7280")
  )
```

::: {.callout-note}
**At a glance:** the week ranged from `r round(summary_stats$low)` F to `r round(summary_stats$high)` F, with `r round(summary_stats$rain, 2)` inches of rain. The strongest hourly wind was `r round(summary_stats$windy)` mph.
:::

## Weather conditions

```{r}
#| label: conditions
#| fig-width: 11
#| fig-height: 3.5

weather |>
  count(date, weather) |>
  ggplot(aes(x = date, y = n, fill = weather)) +
  geom_col(width = 0.75, show.legend = FALSE) +
  geom_text(aes(label = weather), position = position_stack(vjust = 0.5),
            size = 3, color = "#173042", fontface = "bold") +
  scale_x_date(date_breaks = "1 day", date_labels = "%a\n%b %-d") +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "Hourly observations") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title = element_text(color = "#264653", face = "bold"),
    axis.text = element_text(color = "#4a5759")
  )
```

_Data source: [Open-Meteo](https://open-meteo.com/), queried by `plot.R`._
